### 2013-06-02 doc/dissertation/document/chapters/spn_equations/spn_equations.tex
finished draft of solver results section

\subsection{Performance Results}
\label{subsec:spn_comparison_results}
Using the same set of solver parameters used for verification, MCSA
will be directly compared to the production Krylov solvers in terms of
both iterative performance and CPU timing. In addition to the
parameters used for the verification analysis, artificial absorption
was also introduced with a value of 0.2 for the MCSA calculations that
used to collision esitmator as CPU timing performance was improved
without a loss in iterative performance. Adding this absorption did
not change the numerical results of the
calculation. Figure~\ref{fig:spn_comparison_iterations} gives the
average number of linear solver iterations required to converge a
single eigenvalue iteration of the fuel assembly problem as a function
of the number of energy groups.

@@

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/solver_time.pdf}
  \end{center}
  \caption{\textbf{Average CPU time per iteration in seconds for the
      fuel assembly problem as a function of energy groups.}
    \textit{All linear solver iterations over all eigenvalue
      iterations were used to compute the averge. All methods were
      preconditioned with identical ILUT parameters. For MCSA, several
      variations of fixed point iterations and estimators were used.}}
  \label{fig:spn_comparison_time}
\end{figure}

@@

>>>>

\subsection{Performance Results}
\label{subsec:spn_comparison_results}
Using the same set of solver parameters used for verification, MCSA
will be directly compared to the production Krylov solvers in terms of
both iterative performance and CPU timing. In addition to the
parameters used for the verification analysis, artificial absorption
was also introduced with a value of 0.2 for the MCSA calculations that
used the collision esitmator as CPU timing performance was improved
without a loss in iterative performance. Adding this absorption did
not change the numerical results of the
calculation.

Figure~\ref{fig:spn_comparison_iterations} gives the average number of
linear solver iterations required to converge a single eigenvalue
iteration of the fuel assembly problem as a function of the number of
energy groups. In general, the iterative performance of all the
solvers tested is comparable with BiCGStab performing the best. We
expect this because not only are the multigroup $SP_N$ equations
elliptic and positive-definite, but they are also nearly symmetric and
therefore we expect the conjugate gradient-based methods to perform
well. It also important to note that the iterative performance of each
solver is not a strong function of the number of energy groups in the
problem (and therefore the problem size). For MCSA, we are linearly
increasing the number of stochastic histories used to compute the
correction at each iteration as the number of energy groups are
increased and therefore the work performed for the problem is of
$O(N)$. It is also important to note the the reduced domain
approximation parameters, particularly the fill level, were fixed as
the number of energy groups was increased. By fixing the fill level,
we are effectively fixing the amount of information contained in the
composite linear operator available for the Monte Carlo
problem. Limiting this information to 100 entries per row in the
preconditioned system did not appear to have a significant effect on
the iterative performance of the solver. For larger numbers of energy
groups (and therefore DOFs) this may be an issue and this number may
need to be increased to maintain iterative performance. Currently,
memory restrictions arising from the explicit preconditioning strategy
prevent larger energy groups to be analyzed using the appropriate
preconditioning for the fuel assembly problem.

@@

Although iterative performance for MCSA was comparable to that
observed for production Krylov implementations using exactly the same
preconditioning scheme, timing performance is not as
comparable. Figure~\ref{fig:spn_comparison_time} gives the average CPU
time per linear solver iteration required to converge the fuel
assembly problem with all linear solver iterations over all eigenvalue
iterations used to compute the average. The production Krylov solver
implementations are approximately an order of magnitude faster than
the MCSA implementations presented here. We expect these results for
several reasons. First, even when the reduced domain approximation is
used there are O(100) elements in each row of the system and each of
those elements must be processed during a transition in the Monte
Carlo random walk sequence. In addition, although the preconditioned
fixed point iterations used in the MCSA sequence do not use the
composite linear operator but rather a sequence of matrix vector
multiplies to achieve the same effect, they do use the explicitly
extracted inverses of the preconditioning matrices which themselves
are dense, leading to exceedingly slow computation times even in the
fixed point iteration. Second, the implementation presented here has
not been optimized and it is likely several components of the
algorithm can be implemented in a more desireable time complexity. The
production Krylov solvers presented here have enjoyed decades of
professional development and optimization. However, it is most
important to note here that the CPU timing data presented here shows
that all methods have the same time complexity; their differences are
merely the manifestation of a large time constant due to the effects
of the explicit preconditioning scheme. For transport systems where
this is not a problem, the literature explcitly shows that MCSA can be
competetive with these production Krylov methods using CPU time as a
measure \citep{evans_monte_2012}.

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/solver_time.pdf}
  \end{center}
  \caption{\textbf{Average CPU time per linear solver iteration in
      seconds for the fuel assembly problem as a function of energy
      groups.}  \textit{All linear solver iterations over all
      eigenvalue iterations were used to compute the averge. All
      methods were preconditioned with identical ILUT parameters. For
      MCSA, several variations of fixed point iterations and
      estimators were used.}}
  \label{fig:spn_comparison_time}
\end{figure}

Finally, not considered in Figure~\ref{fig:spn_comparison_time} is the
time required to actually form the explicit inverse preconditioner
matrices and the composite operator through matrix-matrix
multiplication. The timing numbers reported were simply to perform the
CMSA iteration procedure with the composite operator already
formed. Figure~\ref{fig:spn_comparison_prec_time} additionally
presents the CPU times for MCSA convergence with the time to form the
preconditioners and the composite linear operator through
matrix-matrix multiplication amortized over all iterations. As is
readily observed including the costs of these operations increases the
MCSA computation time by another order of magnitude.

@@

Based on the results in this section, MCSA shows promise as a
competitive and perhaps even superior method for solutions to the
neutron transport problem discretized with the $SP_N$
approximation. Not only are the correct answers produced when compared
to production linear solvers, but the iterative performance is
comparable to production Krylov methods with identical preconditioning
that would typically be used every-day calculations. From a CPU timing
perspective, the methods yield the same time complexity as a function
of the number of energy groups in the problem when compared to the
production Krylov methods. Here, a large time constant is generated
due to the explicit preconditioning strategy and the generation of
dense linear operators as a result. To improve these results and put
general MCSA schemes into a performance regime where they are
competitive with Krylov methods for neutron transport problems,
significant research will be required to improve upon the explicit
preconditioning shceme presented here.
